\documentclass[a4paper,10pt]{article}
\usepackage{amsmath, amssymb, graphics}
\usepackage{color,cite,graphicx}

\newcommand{\mathsym}[1]{{}}
\newcommand{\unicode}{{}}



%opening
\title{PhD Research Proposal : High Throughput Ultrastructural Phenotype Study of Mouse Femurs}
\author{Kevin Mader}

\begin{document}

\maketitle

\begin{abstract}
The aim of the project is to asses the ultrastructure of bone on a genome-wide scale on the femur bones from 2000 mice. Specifically the vascular network and the osteocyte lacunar system within trabecular and cortical bone will be quantitatively assessed using proven imaging strategies. This experiment will be the assessing the first 200 samples of this group.
\end{abstract}
\tableofcontents
\newpage


\section{Project Overview}
\subsection{Summary}
The goal of the project is to assess the ultrastructure of the femur bones from a set of 2000 mice, carefully bred and raised to provide a genome-wide sampling with minimal variation in environmental factors. Thus the variations in observations will come from genetic differences. Specifically the vascular network and the osteocyte lacunar system within trabecular and cortical bone will be quantitatively assessed using proven imaging strategies. The aim of this particular experiment is to image the first 200 of these bone samples. This experiment will also serve as the first large scale study enabled by the new high-throughput tomograhy system developed at the Tomcat beamline which allows samples to be automatically loaded, aligned, and unloaded without user intervention in quantities of 60 at a time. Significantly reducing the potential for user error, and eliminating wasted time between experiments due to opening, re-mounting, re-aligning, and security checks. This study is in reality the extension of a previously done study at the Institute for Biomechanics looking at the same bone samples and attributing genetic influence at a much lower resolution using a MicroCT scanner with 50x larger pixel size. So while the specific parameters being investigated are different, the methods for genetic analysis will be identical. Additionally the previous study will allow for the parameters in our study to be validated where overlap exists enabling a very easy check on whether the samples were measured, and analyzed correctly. 
\subsection{Biological Background}
Osteoporosis leads to higher risk of bone fracture through decreased bone mass and architectural changes leading to decreased bone quality [3]. While bone mineral density (BMD) is the most important factor in assessing fracture risk, utilizing morphological parameters of cortical bone can significantly increase the predictive power [4,5]. Specifically with cortical bone, BMD has been shown to be a weak factor in predicting mechanical properties, while microarchitectural parameters and porosity measurements are much better indicators for strength [6,7]. The most problematic of the fractures incurred through bone weakness is the proximal femur fracture. Thus the femur bone from mice will be analyzed.
A previous study has shown a strong connection between genetic background and morphology of bone samples, but has only done so on the scale of a few animals [1]. This study, however, is insufficient to determine the links between genetic background and bone ultrastructure. A genome-wide set of inbred mice has been created at The Jackson Laboratory (Bay Harbor, Maine, USA). The mice were inbred for many generations to ensure the genetic purity of each sample. Furthermore the mice were raised in an identical environment to minimize the effect environmental factors. The femurs have been extracted from these mice and reside at the Institute for Biomechanics at the ETH.  
We believe this first of its kind study will lead to revolutions in the understanding of genetic regulation of bone structure and bring more complete information to the studies of mechanism in genetic disorders and diseases in bone. Furthermore the information will provide a quantitative backing to re-examine previously purely epidemiological studies revealing new possibilities for treatment and prevention. Against osteoporosis we expect to lay the foundation for identifying new targets for treatment, and particularly at risk populations and provide the tools to much more accurately estimate the efficacy of treatments based on genetic background.


\begin{itemize}
 	\item Canal Interest
	\begin{itemize}
		\item Establish baseline, homeostatic parameters for canals and canal networks
		\item Correlate the variation in these parameters to macroscopic observed phenomenon (bone strength, bone density, …)
	\end{itemize}
	\item Lacunae Interest
	\begin{itemize}
		\item Establish baseline, homeostatic parameters for lacunae
		\item Correlate the variation in these parameters to macroscopic observed phenomenon (bone strength, bone density, …)
	\end{itemize}
	\item Outlook
	\begin{itemize}
		\item Some myleoma cells embed themselves in bone-matrix and become immortal and untreatable through techniques, further understanding of the behavior of osteocytes and how they interact with the matrix could be useful in developing new treatments to differentiate between these cells and carcinogenic cells.
	\end{itemize}
\end{itemize}








\newpage
\section{Pre-measuement Tasks}
\include{PreMeasurement}

\newpage
\section{Measurement}
\include{Measurement}

\newpage
\section{Post-Measurement Analysis}
\include{PostMeasurement}


\section{Conclusion}
As seen in the above report a majority of the development work for the project should be done on or around the completion of the first beamtime, with only minor additional tasks left to be performed over the following months. The extra-time will allow ample time to conduct all the measurements and remeasuring in the event of other problems. After all the samples are measured which would ideally be completed by the end of year 2, there will be plenty of time to analyze the results, publish and possibly begin investigatory studies into other aspects of bone structure for example looking into mineralization density and variations and the genetic influences there. The methods and tools developed to do this study will enable future studies of such scale to be done much quicker, and establish a standard pipeline where further genetic and other large sample number studies could be done in several beamtimes with minimal user intervention and without the need to develop new tools.
\end{document}
